# Code & Data {#sec-code}

This chapter exercises figures, tables, code blocks, code windows (with auto-filename), script injection, and tabsets.

## Figure {#sec-figure}

@fig-penguins comes from **ggplot2** on the `palmerpenguins` dataset, with the shared knitr options from `_extension.yml` (`echo: false`, `message: false`, `warning: false`).
The HTML output wraps figures in a lightbox; click the image to enlarge.

```{r}
#| label: fig-penguins
#| fig-cap: "Bill length vs flipper length by species (palmerpenguins)."
#| fig-width: 6
#| fig-height: 4

library(ggplot2)

ggplot(
  palmerpenguins::penguins,
  aes(x = flipper_length_mm, y = bill_length_mm, colour = species)
) +
  geom_point(size = 3, alpha = 0.85) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.6) +
  labs(
    x = "Flipper length (mm)",
    y = "Bill length (mm)",
    colour = "Species"
  ) +
  theme_minimal(base_size = 12)
```

## Table {#sec-table}

`tbl-cap-location: top` (set in `_extension.yml`) places the caption above the table.
`tbl-title: "Table"` is language-neutral and set extension-wide.

```{r}
#| label: tbl-species
#| tbl-cap: "Penguin counts by species (palmerpenguins)."
#| echo: true
#| code-fold: false

library(dplyr)

palmerpenguins::penguins |>
  count(species) |>
  rename(Species = species, N = n)
```

## Code block

The chunk below sets `echo: true` (overriding the `_extension.yml` default) to exercise syntax highlighting, code folding, and the copy-on-hover button.

```{r}
#| echo: true
#| eval: false

library(dplyr)

palmerpenguins::penguins |>
  filter(!is.na(body_mass_g)) |>
  group_by(species, island) |>
  summarise(
    n = n(),
    mean_mass_g = mean(body_mass_g),
    .groups = "drop"
  ) |>
  arrange(desc(mean_mass_g))
```

## Code window {#sec-code-window}

The blocks below exercise the [`mcanouil/quarto-code-window`](https://github.com/mcanouil/quarto-code-window) filter.
`style: default` gives a plain filename header; `auto-filename: true` shows the language name for blocks without an explicit `filename=`.

### Auto-generated filename (language name)

```python
def greet(name: str) -> str:
    return f"Hello, {name}!"
```

```r
penguins |>
  dplyr::count(species)
```

```bash
quarto add hebstr/quarto-hebstr-book --no-prompt
quarto render
```

## Script fold {#sec-script}

The block below uses `code-fold: true` to produce a collapsible `<details>` element with the filename as summary.
In standalone documents, the `{{</* script */>}}` shortcode (in-tree `filters/add-code-files.lua`) injects external script files with the same visual result.

{{< script scripts/demo_penguins.R >}}

## Tabset {#sec-tabs}

Tabs use the `.panel-tabset` div class; `tab-background` from `theme-light.scss` / `theme-dark.scss` sets the panel fill.

::: {.panel-tabset}

## Adelie

```{r}
#| echo: true
#| code-fold: false
palmerpenguins::penguins |>
  dplyr::filter(species == "Adelie") |>
  dplyr::select(island, bill_length_mm, flipper_length_mm, body_mass_g) |>
  head(5)
```

## Chinstrap

```{r}
#| echo: true
#| code-fold: false
palmerpenguins::penguins |>
  dplyr::filter(species == "Chinstrap") |>
  dplyr::select(island, bill_length_mm, flipper_length_mm, body_mass_g) |>
  head(5)
```

## Gentoo

```{r}
#| echo: true
#| code-fold: false
palmerpenguins::penguins |>
  dplyr::filter(species == "Gentoo") |>
  dplyr::select(island, bill_length_mm, flipper_length_mm, body_mass_g) |>
  head(5)
```

:::
